### 2022-08-08 sandp23/background.tex
Adding mods to intro

\section{Background}
\subsection{Program Partitioning}
\label{subsec:background:programpart}