### 2023-01-13 sandp23/limitations.tex
Adding implememtation

\section{Limitations}
\label{sec:limitations}
We cannot handling function pointer calls to tainted function.