### 2023-01-15 sandp23/tables/datasettable.tex
Adding all

{
\footnotesize
\begin{table}[]
\begin{tabular}{c|c|c|r}
\toprule
\textbf{ID} & \textbf{Program} & \textbf{Description}          & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Size\\ (SLoc)\end{tabular}}} \\ 
\midrule
\rowcolor{black!15} 1           & ProFTPD          & High performance FTP Server   & XXX K                                                                               \\ 
2           & MicroHTTPD       & Simple HTTPD Server           & YY K                                                                                \\ 
\rowcolor{black!15} 3           & UFTPD            & UDP based FTP Server          & ZZZ K                                                                               \\ 
4           & png2pnm          & Program to convert png to pnm & UUU K                                                                               \\ 
\bottomrule
\end{tabular}
\caption{Evaluation Dataset.}
\label{table:dataset}
\end{table}
}