### 2023-01-17 sandp23/tables/datasettable.tex

{
\footnotesize
\begin{table}[]
\begin{tabular}{c|c|c|r}
\toprule
\textbf{ID} & \textbf{Program} & \textbf{Description}          & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Size\\ (SLoc)\end{tabular}}} \\ 
\midrule
\rowcolor{black!15} 1           & ProFTPD          & High performance FTP Server   & 556 K                                                                               \\ 
2           & MicroHTTPD       & Simple HTTPD Server           & 122 K                                                                                \\ 
\rowcolor{black!15} 3           & UFTPD            & UDP based FTP Server          & 3 K                                                                               \\ 
4           & LibPNG          & Program to convert between png and pnm & 76 K                                                                               \\ 
\rowcolor{black!15}5           & TinyBigNum          & Multiple-precision integer implementation & 1.6 K                                                                               \\ 
6           & Parsons          & JSON parsing library & 3.1 K                                                                               \\

\bottomrule
\end{tabular}
\caption{Evaluation Dataset.}
\label{table:dataset}
\end{table}
}