### 2023-12-01 sandp23/figures/newusecase.tex

\begin{figure}
\begin{center}
\begin{tikzpicture}[scale=0.75]
%\node[text width=3cm] at (1.5,-0.5) 
%    {some text spanning three};
%\draw[step=5cm,gray,very thin] (0,0) grid (5,5);
\fill[taintcolor] (2.5,2.5) rectangle (5,5);
\fill[green!10!white] (0,2.5) rectangle (2.5,5);
\fill[green!40!white] (2.5,0) rectangle (5,2.5);
\fill[green!40!white] (0,0) rectangle (2.5,2.5);

\Text[x=3.8cm,y=4.5cm,fontsize=\scriptsize]{\bf High Risk and}
\Text[x=3.8cm,y=4.2cm,fontsize=\scriptsize]{\bf Hard to Convert}
\Text[x=3.8cm,y=3.9cm,fontsize=\scriptsize]{\bf Functions.}
\Text[x=3.8cm,y=3.6cm,fontsize=\tiny]{(Isolated in}
\Text[x=3.8cm,y=3.3cm,fontsize=\tiny]{the Sandbox)}

\Text[x=1.2cm,y=4.5cm,fontsize=\scriptsize]{\bf Low Risk and}
\Text[x=1.2cm,y=4.2cm,fontsize=\scriptsize]{\bf Hard to Convert.}
\Text[x=1.2cm,y=3.9cm,fontsize=\scriptsize]{\bf Functions.}
\Text[x=1.2cm,y=3.6cm,fontsize=\tiny]{(Partially annotated}
\Text[x=1.3cm,y=3.3cm,fontsize=\tiny]{and execute in c-region)}
\Text[x=1.2cm,y=2.9cm,fontsize=\tiny]{}
%\Text[x=1.2cm,y=4.3cm,fontsize=\tiny]{(Will be partially converted)}

\Text[x=2.3cm,y=1.8cm,fontsize=\scriptsize]{\bf Easy to Convert Functions.}
\Text[x=2.3cm,y=1.5cm,fontsize=\tiny]{(Will be fully converted to Checked C}
\Text[x=2.3cm,y=1.2cm,fontsize=\tiny]{with all necessary annotations}
\Text[x=2.3cm,y=0.9cm,fontsize=\tiny]{ and execute in c-region)}

%\Text[x=3.8cm,y=2.1cm,fontsize=\tiny]{Retrofitting}
%\Text[x=3.8cm,y=1.8cm,fontsize=\tiny]{Techniques.}
%\Text[x=3.8cm,y=1.5cm,fontsize=\tiny]{(ASAN, CCured,}
%\Text[x=3.8cm,y=1.3cm,fontsize=\tiny]{etc)}

%\Text[x=0.9cm,y=0.6cm,fontsize=\tiny]{\bf \faFlagCheckered}

%\Text[x=2.0cm,y=1.4cm,fontsize=\tiny]{\bf \footnotesize iC3C}

%\draw [{Implies}-,double,line width=1pt] (1.0,0.8) -- (1.9,2.1);

\draw[-Stealth,thick] (0,0) -- (5.1,0) node[anchor=north west] {\bf \tiny high};
\Text[x=0cm,y=-0.2cm,fontsize=\tiny]{\bf least}
\Text[x=2.6cm,y=-0.2cm,fontsize=\footnotesize]{Functions' Risk Level}
\Text[x=-0.2,y=2.6,rotation=90,fontsize=\footnotesize]{Conversion effort}
\draw[-Stealth,thick] (0,0) -- (0,5.1) node[anchor=south east] {\bf \tiny high};
\end{tikzpicture}
\caption{Flexibile partioning using~\systemname{}.}
\label{fig:mixedmodesupport}
\end{center}
\end{figure}